---
title: "P8451 Assignment 1"
output: html_document
date: "2023-01-18"
---

```{r}
library(tidyverse)
library(psych)
library(dplyr)
```

## Importing and Cleaning Data

```{r}
bcdata <- read_csv("data/bcdata_Assignment1.csv") %>%
  janitor::clean_names() %>%
  drop_na()
```

## Question 1

Construct a table providing summaries of the quantitative features of the dataset. Summaries should include the mean, median, minimum value, and maximum value. 

```{r}
  describe(bcdata) %>%
  knitr::kable()
```

## Question 2

Recode BMI into the WHO-defined categories below:

* Severely underweight - BMI less than 16.5 kg/m^2

* Underweight - BMI under 18.5 kg/m^2

* Normal weight - BMI greater than or equal to 18.5 to 24.9 kg/m^2

* Overweight - BMI greater than or equal to 25 to 29.9 kg/m^2 

* Obesity class I - BMI 30 to 34.9 kg/m^2 

* Obesity class II - BMI 35 to 39.9 kg/m^2

* Obesity class III - BMI greater than or equal to 40 kg/m^2 

```{r}
bcdata$bmi_categories <- as.factor( ifelse (bcdata$bmi < 16.5, 'Severely Underweight',
                                    ifelse (bcdata$bmi < 18.5, 'Underweight',
                                    ifelse (bcdata$bmi < 24.9, 'Normal weight',
                                    ifelse (bcdata$bmi < 29.9, 'Overweight',
                                    ifelse (bcdata$bmi < 34.9, 'Obesity class I',
                                    ifelse (bcdata$bmi < 39.9, 'Obesity class II',
                                    ifelse (bcdata$bmi < 40, 'Obesity class III'))))))))

bcdata %>%
  group_by(bmi_categories) %>%
  summarise(n_obs = n()) %>%
  knitr::kable()
```

## Question 3 

Create a bar chart showing the proportion of breast cancer cases and controls within each BMI category. (That is, the graph should clearly show what % of each BMI category are cases and what % are controls)

```{r}
bcdata = 
  bcdata %>%
  mutate(
    classification = recode(classification ,
                            "1" = "Control",
                            "2" = "Case")
  ) 

bcdata = 
  bcdata %>%
  group_by(bmi_categories) %>%
  mutate(count = n()) %>%
  mutate(bmi_categories = fct_reorder(bmi_categories, count))

ggplot(data = bcdata, (aes(x = reorder(bmi_categories, count), fill = classification))) +
  geom_bar() +
  labs(
    title = "Breast cancer case and control distribution within BMI categories",
    x = "BMI categories") +
  theme(legend.position = "bottom") +
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 10))
```










