---
title: "P8451 Assignment 1"
output: html_document
date: "2023-01-18"
---

```{r}
library(tidyverse)
library(psych)
library(dplyr)
```

## Importing and Cleaning Data

```{r}
bcdata <- read_csv("data/bcdata_Assignment1.csv") %>%
  janitor::clean_names() %>%
  drop_na()
```

## Question 1

Construct a table providing summaries of the quantitative features of the dataset. Summaries should include the mean, median, minimum value, and maximum value. 

```{r}
  describe(bcdata) %>%
  knitr::kable()
```

## Question 2

Recode BMI into the WHO-defined categories below:

* Severely underweight - BMI less than 16.5 kg/m^2

* Underweight - BMI under 18.5 kg/m^2

* Normal weight - BMI greater than or equal to 18.5 to 24.9 kg/m^2

* Overweight - BMI greater than or equal to 25 to 29.9 kg/m^2 

* Obesity class I - BMI 30 to 34.9 kg/m^2 

* Obesity class II - BMI 35 to 39.9 kg/m^2

* Obesity class III - BMI greater than or equal to 40 kg/m^2 

```{r}
bcdata$bmi_categories <- as.factor( ifelse (bcdata$bmi < 16.5, 'Severely Underweight',
                                    ifelse (bcdata$bmi < 18.5, 'Underweight',
                                    ifelse (bcdata$bmi < 25, 'Normal weight',
                                    ifelse (bcdata$bmi < 30, 'Overweight',
                                    ifelse (bcdata$bmi < 35, 'Obesity class I',
                                    ifelse (bcdata$bmi < 40, 'Obesity class II',
                                    'Obesity class III')))))))

bcdata %>%
  group_by(bmi_categories) %>%
  summarise(n_obs = n()) %>%
  knitr::kable()
```

## Question 3 

Create a bar chart showing the proportion of breast cancer cases and controls within each BMI category. (That is, the graph should clearly show what % of each BMI category are cases and what % are controls)

```{r}
bcdata %>%
  group_by(bmi_categories) %>%
  mutate(n = n(),
         cases = sum(classification - 1),
         proportion_cases = cases / n) %>%
  ggplot(aes(x = reorder(bmi_categories, proportion_cases), fill = factor(classification))) +
  geom_bar(position = "fill") +
  labs(
    title = "Breast cancer case and control distribution within BMI categories",
    x = "BMI categories",
    y = "Proportion") +
  theme(legend.position = "bottom") +
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 10)) +
  scale_fill_discrete(labels = c("Control", "Case"))
```

## Question 4

Construct a basic regression model using breast cancer classification as the outcome and glucose, HOMA-IR, leptin, BMI (continuous) and age as the independent variables. Fill in the beta estimate and 95% confidence interval associated with a 1-unit change in HOMA-IR

```{r}
bcdata = 
  bcdata %>%
  mutate(
    classification = recode(classification ,
                            "1" = 0,
                            "2" = 1)
  ) 

fit_logistic = 
  bcdata %>%
  glm(classification ~ glucose + homa + leptin + bmi + age, data = ., family = binomial()) %>%
  broom::tidy()

fit_logistic %>%
  mutate(CI_lower = estimate - 1.96*std.error,
         CI_upper = estimate + 1.96*std.error) %>%
  select(term, estimate, CI_lower, CI_upper, p.value) %>%
  knitr::kable(digits = 3)
```

The beta estimate associated with a 1-unit change in HOMA-IR is 0.274. The 95% CI for this beta estimate is (-0.063, 0.611).

## Question 5 

Constuct a linear regression model using insulin as the outcome and BMI (continuous), age, and glucose as the independent variables. Fill in the beta estimate and 95% confidence interval associated with a 1-unit change in age.

```{r}
fit_linear = 
  lm(insulin ~ bmi + age + glucose, data = bcdata)

fit_linear %>%
  broom::tidy() %>%
  mutate(CI_lower = estimate - 1.96*std.error,
         CI_upper = estimate + 1.96*std.error) %>%
  select(term, estimate, CI_lower, CI_upper, p.value) %>%
  knitr::kable(digits = 3)
```

The beta estimate associated with a 1-unit change in age is -0.054. The 95% CI for this beta estimate is (-0.156, 0.048). 














